\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants shape their interactions with pollinators

### 2023-02-24 Manuscript/Lanuza_et_al_2022.Rmd
New two paragraphs on the discussion, much more extended now

Overall, plant reproductive traits explained interaction patterns among and within pollinator guilds suggesting that plant reproductive strategies partly mediate these associations across spatial large scales. For example, we found that all pollinator guilds interacted more frequently with species with large floral displays and low ability to self pollinate, which aligns with previous findings [@hegland2005; @lazaro2013; @kaiser2014]. As predicted by evolutionary life-history theory [@stebbins1970], larger investment in floral display and rewards to attract pollinators occurs when pollinator dependence is high. Indeed, the possible range of reproductive strategies in flowering plants is constrained by the cost of reproductive organs, trait associations and selective pressures exerted by different pollinator guilds [@galen1999]. For instance, smaller flowers are visited less overall, but they are inexpensive to produce (e.g., have a low ovule number) and low visitation rate at the individual flower level can be compensated by overall visitation at the plant level by producing many flowers resulting in a larger quantity of small fruits [@kettle2011]. In addition, we found that a substantial number of species with high autonomous selfing had large styles and were associated with a high number of interactions with bees. These species seem to rely on animal pollinators as indicated by their high investment in sizable styles that can help prevent pollen competition [@ramesha2011; @ashman2013; @lanuza2021] while having reproductive assurance in their absence. For instance, the annual herb _Scabiosa atropurpurea_ has flowers grouped in clusters with relatively long styles and a single ovule per flower that is likely to be fertilised with a single legitimate floral visit but avoids reproductive failure with autonomous self pollination. Similarly, zygomorphic flowers are specialised floral structures that promote precise pollen transfer [@sargent2004; @gong2009] but they were associated with high levels of autogamy. Our findings are consistent with the idea that reproductive assurance is not rare within specialised pollination systems [@fenster2007] and highlight the need to explore the timing (e.g., early versus delayed) and selfing mechanisms [see @goodwillie2018] in relation to other reproductive traits to better understand reproductive trait correlations and plant-pollinator interactions.

Because floral traits can be associated with their animal pollinators to optimise pollen transfer, the correlation between distinct suites of reproductive traits and pollinator guilds could indicate the existence of pollination syndromes [@fenster2004; @faegri2013; @rosas2014; @dellinger2020]. For instance, bees had a higher number of interacting partners on plant species with large flowers but interacted more frequently with species with small flowers while syrphid flies showed the opposite trend. This could be reflective of differences in the ecology and behaviour between these pollinator guilds [@rader2020; @foldesi2021] that ultimately exert different selective pressures, leading divergent plant reproductive strategies. In any case, it is worth noting that local factors (e.g., species relative abundances or phenologies) likely contribute to the variability in the association between the plant reproductive spectrum and pollinators [@vazquez2007; @encinas2012; @bartomeus2016]. Taking both aspects together, we suggest that while plant reproductive traits do shape their interactions with pollinators, pollinator syndromes at a macroecological scale are not distinct compartmentalised items, but form part of a multifaceted context dependent continuum.

>>>>

Overall, the main plant reproductive axes were insufficient to fully capture plant-pollinator associations. However, these axes helped explain, to some extent, interaction patterns among and within pollinator guilds, indicating that they partly mediate these associations across large spatial scales. For example, we found that all pollinator guilds interacted more frequently with species with large floral displays and low ability to self pollinate, which aligns with previous findings [@hegland2005; @lazaro2013; @kaiser2014] and predictions from evolutionary life-history theory [@stebbins1970]. These predictions suggest that larger investment in floral display and rewards to attract pollinators occurs when dependence on pollinators is high. Although pollinators play a major role in shaping plant reproductive strategies, these different strategies are also constrained by the cost of reproductive organs and trait correlations [@galen1999]. For instance, smaller flowers tend to receive fewer visits than larger flowers, but small flowers are inexpensive to produce (e.g., have few ovules). Therefore, low  per-flower visitation can be compensated by overall plant-level visitation through producing many small flowers that develop into numerous small fruits [@kettle2011]. Our results also indicate that bees interact more with plants that have higher levels of autonomous selfing and large styles, which can help with a more efficient pollen transfer and prevent pollen competition [@ashman2013; @tong2016; @lanuza2021]. Similarly, the specialised floral structures of zygomorphic flowers promote precise pollen transfer [@sargent2004; @gong2009], although we found that they were associated with high levels of autogamy. The annual herb _Lathyrus clymenum_ is a clear example of this pollinator specialization - selfing strategy that may get “the best of both worlds” [@lloyd1992; @goodwillie2018]. It has showy zygomorphic flowers with relatively long styles and few ovules per flower that are likely fertilised with a single legitimate floral visit but can avoid reproductive failure in the absence of pollinators through autonomous self pollination. Our findings are consistent with the hypothesis that selfing is not rare within specialised pollination systems [@fenster2007] and emphasise the need to consider different reproductive assurance mechanisms [e.g., early versus delayed selfing\; @goodwillie2018] to better understand plant reproductive trait correlations and plant-pollinator interactions.

Because flowering plants can adapt to their animal pollinators to optimise pollen transfer, the correlation between distinct suites of plant reproductive traits and pollinator guilds could indicate the existence of pollination syndromes [@fenster2004; @faegri2013; @rosas2014; @dellinger2020]. However, our work shows a weak association between the different plant reproductive strategies and pollinator guilds, which supports previous findings from a similar pollinator guild classification and spatial scale [@ollerton2009]. Although our findings contrast with a vast amount of literature on pollinator syndromes dating back over half a century [@faegri2013; @fenster2004], the ecological scale of our study differs from the majority of floral syndrome studies that rarely consider plant-pollinator associations beyond the community scale [ca. 3% of studies\; @dellinger2020]. We argue that the predictive power of floral syndromes gets blurred when increasing the number of species and scale, reinforcing the view that strong specialisation may be the exception rather than the norm in plant-pollinator interactions [@jordano1987; @waser1996]. Further, there are several factors that might have influenced our findings, (e.g., the suite of reproductive traits and type of pollinator guilds we included). For instance, including other key reproductive traits, such as floral scent or flower colour, could help to improve predictions of plant-pollinator interactions. In addition, our use of broad taxonomic pollinator guild groupings may have obscured  finer scale patterns that occur at lower taxonomic ranks (e.g., genus or species level). The fact that our study only accounts for insect pollinators may have also influenced the observed associations between floral traits and pollinator guilds in comparison with other studies, as some other guilds (e.g., hummingbirds or bats) generally display strong trait-matching [@wilson2004; @lagomarsino2017]. Regardless, it is worth noting that local factors (e.g., species relative abundances, phenologies or intraspecific variation) likely contribute to the variability in interactions between the plant reproductive spectrum and pollinators [@vazquez2007; @encinas2012; @bartomeus2016; @arroyo2023]. We suggest that while plant reproductive traits do influence which pollinators they interact with, at a macroecological scale, pollination syndromes possibly take the form of a multifaceted, context dependent continuum [e.g., ornithophilous versus entomophilous pollination of red flowers\; @leon2022].

### 2023-02-24 Manuscript/Lanuza_et_al_2022.tex
New two paragraphs on the discussion, much more extended now

Overall, plant reproductive traits explained interaction patterns among and within pollinator guilds suggesting that plant reproductive strategies partly mediate these associations across spatial large scales. For example, we found that all pollinator guilds interacted more frequently with species with large floral displays and low ability to self pollinate, which aligns with previous findings (\protect\hyperlink{ref-hegland2005}{Hegland \& Totland, 2005}; \protect\hyperlink{ref-kaiser2014}{Kaiser-Bunbury et al., 2014}; \protect\hyperlink{ref-lazaro2013}{Lázaro et al., 2013}). As predicted by evolutionary life-history theory (\protect\hyperlink{ref-stebbins1970}{Stebbins, 1970}), larger investment in floral display and rewards to attract pollinators occurs when pollinator dependence is high. Indeed, the possible range of reproductive strategies in flowering plants is constrained by the cost of reproductive organs, trait associations and selective pressures exerted by different pollinator guilds (\protect\hyperlink{ref-galen1999}{Galen, 1999}). For instance, smaller flowers are visited less overall, but they are inexpensive to produce (e.g., have a low ovule number) and low visitation rate at the individual flower level can be compensated by overall visitation at the plant level by producing many flowers resulting in a larger quantity of small fruits (\protect\hyperlink{ref-kettle2011}{Kettle et al., 2011}). In addition, we found that a substantial number of species with high autonomous selfing had large styles and were associated with a high number of interactions with bees. These species seem to rely on animal pollinators as indicated by their high investment in sizable styles that can help prevent pollen competition (\protect\hyperlink{ref-ashman2013}{Ashman \& Arceo-Gómez, 2013}; \protect\hyperlink{ref-lanuza2021}{Lanuza et al., 2021}; \protect\hyperlink{ref-ramesha2011}{Ramesha et al., 2011}) while having reproductive assurance in their absence. For instance, the annual herb \emph{Scabiosa atropurpurea} has flowers grouped in clusters with relatively long styles and a single ovule per flower that is likely to be fertilised with a single legitimate floral visit but avoids reproductive failure with autonomous self pollination. Similarly, zygomorphic flowers are specialised floral structures that promote precise pollen transfer (\protect\hyperlink{ref-gong2009}{Gong \& Huang, 2009}; \protect\hyperlink{ref-sargent2004}{Sargent, 2004}) but they were associated with high levels of autogamy. Our findings are consistent with the idea that reproductive assurance is not rare within specialised pollination systems (\protect\hyperlink{ref-fenster2007}{Fenster \& Marten-Rodriguez, 2007}) and highlight the need to explore the timing (e.g., early versus delayed) and selfing mechanisms (see \protect\hyperlink{ref-goodwillie2018}{Goodwillie \& Weber, 2018}) in relation to other reproductive traits to better understand reproductive trait correlations and plant-pollinator interactions.

Because floral traits can be associated with their animal pollinators to optimise pollen transfer, the correlation between distinct suites of reproductive traits and pollinator guilds could indicate the existence of pollination syndromes (\protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}; \protect\hyperlink{ref-faegri2013}{Faegri \& Van Der Pijl, 2013}; \protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}; \protect\hyperlink{ref-rosas2014}{Rosas-Guerrero et al., 2014}). For instance, bees had a higher number of interacting partners on plant species with large flowers but interacted more frequently with species with small flowers while syrphid flies showed the opposite trend. This could be reflective of differences in the ecology and behaviour between these pollinator guilds (\protect\hyperlink{ref-foldesi2021}{Földesi et al., 2021}; \protect\hyperlink{ref-rader2020}{Rader et al., 2020}) that ultimately exert different selective pressures, leading divergent plant reproductive strategies. In any case, it is worth noting that local factors (e.g., species relative abundances or phenologies) likely contribute to the variability in the association between the plant reproductive spectrum and pollinators (\protect\hyperlink{ref-bartomeus2016}{Bartomeus et al., 2016}; \protect\hyperlink{ref-encinas2012}{Encinas-Viso et al., 2012}; \protect\hyperlink{ref-vazquez2007}{Vázquez et al., 2007}). Taking both aspects together, we suggest that while plant reproductive traits do shape their interactions with pollinators, pollinator syndromes at a macroecological scale are not distinct compartmentalised items, but form part of a multifaceted context dependent continuum.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-foldesi2021}{}}%
Földesi, R., Howlett, B. G., Grass, I., \& Batáry, P. (2021). Larger pollinators deposit more pollen on stigmas across multiple plant species---a meta-analysis. \emph{J. Appl. Ecol.}, \emph{58}(4), 699--707. \url{https://doi.org/10.1111/1365-2664.13798}

@@

@@

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-rader2020}{}}%
Rader, R., Cunningham, S., Howlett, B., \& Inouye, D. (2020). Non-bee insects as visitors and pollinators of crops: Biology, ecology, and management. \emph{Annu. Rev. Entomol.}, \emph{65}, 391--407. \url{https://doi.org/10.1146/annurev-ento-011019-025055}

\leavevmode\vadjust pre{\hypertarget{ref-ramesha2011}{}}%
Ramesha, B., Yetish, M., Ravikanth, G., Ganeshaiah, K., Ghazoul, J., \& Shaanker, R. U. (2011). Stylish lengths: Mate choice in flowers. \emph{Journal of Biosciences}, \emph{36}(2), 229--234. \url{https://doi.org/10.1007/s12038-011-9057-6}

@@

@@

>>>>

Overall, the main plant reproductive axes were insufficient to fully capture plant-pollinator associations. However, these axes helped explain, to some extent, interaction patterns among and within pollinator guilds, indicating that they partly mediate these associations across large spatial scales. For example, we found that all pollinator guilds interacted more frequently with species with large floral displays and low ability to self pollinate, which aligns with previous findings (\protect\hyperlink{ref-hegland2005}{Hegland \& Totland, 2005}; \protect\hyperlink{ref-kaiser2014}{Kaiser-Bunbury et al., 2014}; \protect\hyperlink{ref-lazaro2013}{Lázaro et al., 2013}) and predictions from evolutionary life-history theory (\protect\hyperlink{ref-stebbins1970}{Stebbins, 1970}). These predictions suggest that larger investment in floral display and rewards to attract pollinators occurs when dependence on pollinators is high. Although pollinators play a major role in shaping plant reproductive strategies, these different strategies are also constrained by the cost of reproductive organs and trait correlations (\protect\hyperlink{ref-galen1999}{Galen, 1999}). For instance, smaller flowers tend to receive fewer visits than larger flowers, but small flowers are inexpensive to produce (e.g., have few ovules). Therefore, low per-flower visitation can be compensated by overall plant-level visitation through producing many small flowers that develop into numerous small fruits (\protect\hyperlink{ref-kettle2011}{Kettle et al., 2011}). Our results also indicate that bees interact more with plants that have higher levels of autonomous selfing and large styles, which can help with a more efficient pollen transfer and prevent pollen competition (\protect\hyperlink{ref-ashman2013}{Ashman \& Arceo-Gómez, 2013}; \protect\hyperlink{ref-lanuza2021}{Lanuza et al., 2021}; \protect\hyperlink{ref-tong2016}{Tong \& Huang, 2016}). Similarly, the specialised floral structures of zygomorphic flowers promote precise pollen transfer (\protect\hyperlink{ref-gong2009}{Gong \& Huang, 2009}; \protect\hyperlink{ref-sargent2004}{Sargent, 2004}), although we found that they were associated with high levels of autogamy. The annual herb \emph{Lathyrus clymenum} is a clear example of this pollinator specialization - selfing strategy that may get ``the best of both worlds'' (\protect\hyperlink{ref-goodwillie2018}{Goodwillie \& Weber, 2018}; \protect\hyperlink{ref-lloyd1992}{Lloyd \& Schoen, 1992}). It has showy zygomorphic flowers with relatively long styles and few ovules per flower that are likely fertilised with a single legitimate floral visit but can avoid reproductive failure in the absence of pollinators through autonomous self pollination. Our findings are consistent with the hypothesis that selfing is not rare within specialised pollination systems (\protect\hyperlink{ref-fenster2007}{Fenster \& Marten-Rodriguez, 2007}) and emphasise the need to consider different reproductive assurance mechanisms (e.g., early versus delayed selfing; \protect\hyperlink{ref-goodwillie2018}{Goodwillie \& Weber, 2018}) to better understand plant reproductive trait correlations and plant-pollinator interactions.

Because flowering plants can adapt to their animal pollinators to optimise pollen transfer, the correlation between distinct suites of plant reproductive traits and pollinator guilds could indicate the existence of pollination syndromes (\protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}; \protect\hyperlink{ref-faegri2013}{Faegri \& Van Der Pijl, 2013}; \protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}; \protect\hyperlink{ref-rosas2014}{Rosas-Guerrero et al., 2014}). However, our work shows a weak association between the different plant reproductive strategies and pollinator guilds, which supports previous findings from a similar pollinator guild classification and spatial scale (\protect\hyperlink{ref-ollerton2009}{Ollerton et al., 2009}). Although our findings contrast with a vast amount of literature on pollinator syndromes dating back over half a century (\protect\hyperlink{ref-faegri2013}{Faegri \& Van Der Pijl, 2013}; \protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}), the ecological scale of our study differs from the majority of floral syndrome studies that rarely consider plant-pollinator associations beyond the community scale (ca. 3\% of studies; \protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}). We argue that the predictive power of floral syndromes gets blurred when increasing the number of species and scale, reinforcing the view that strong specialisation may be the exception rather than the norm in plant-pollinator interactions (\protect\hyperlink{ref-jordano1987}{Jordano, 1987}; \protect\hyperlink{ref-waser1996}{Waser et al., 1996}). Further, there are several factors that might have influenced our findings, (e.g., the suite of reproductive traits and type of pollinator guilds we included). For instance, including other key reproductive traits, such as floral scent or flower colour, could help to improve predictions of plant-pollinator interactions. In addition, our use of broad taxonomic pollinator guild groupings may have obscured finer scale patterns that occur at lower taxonomic ranks (e.g., genus or species level). The fact that our study only accounts for insect pollinators may have also influenced the observed associations between floral traits and pollinator guilds in comparison with other studies, as some other guilds (e.g., hummingbirds or bats) generally display strong trait-matching (\protect\hyperlink{ref-lagomarsino2017}{Lagomarsino et al., 2017}; \protect\hyperlink{ref-wilson2004}{Wilson et al., 2004}). Regardless, it is worth noting that local factors (e.g., species relative abundances, phenologies or intraspecific variation) likely contribute to the variability in interactions between the plant reproductive spectrum and pollinators (\protect\hyperlink{ref-arroyo2023}{Arroyo-Correa et al., 2023}; \protect\hyperlink{ref-bartomeus2016}{Bartomeus et al., 2016}; \protect\hyperlink{ref-encinas2012}{Encinas-Viso et al., 2012}; \protect\hyperlink{ref-vazquez2007}{Vázquez et al., 2007}). We suggest that while plant reproductive traits do influence which pollinators they interact with, at a macroecological scale, pollination syndromes possibly take the form of a multifaceted, context dependent continuum (e.g., ornithophilous versus entomophilous pollination of red flowers; \protect\hyperlink{ref-leon2022}{León-Osper \& Narbona, 2022}).

@@

\leavevmode\vadjust pre{\hypertarget{ref-arroyo2023}{}}%
Arroyo-Correa, B., Jordano, P., \& Bartomeus, I. (2023). Intraspecific variation in species interactions promotes the feasibility of mutualistic assemblages. \emph{Ecology Letters}. \url{https://doi.org/10.1111/ele.14163}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-jordano1987}{}}%
Jordano, P. (1987). Patterns of mutualistic interactions in pollination and seed dispersal: Connectance, dependence asymmetries, and coevolution. \emph{The American Naturalist}, \emph{129}(5), 657--677. \url{https://doi.org/10.1086/284665}

@@

\leavevmode\vadjust pre{\hypertarget{ref-lagomarsino2017}{}}%
Lagomarsino, L. P., Forrestel, E. J., Muchhala, N., \& Davis, C. C. (2017). Repeated evolution of vertebrate pollination syndromes in a recently diverged andean plant clade. \emph{Evolution}, \emph{71}(8), 1970--1985.

@@

\leavevmode\vadjust pre{\hypertarget{ref-leon2022}{}}%
León-Osper, M., \& Narbona, E. (2022). Unravelling the mystery of red flowers in the mediterranean basin: How to be conspicuous in a place dominated by hymenopteran pollinators. \emph{Functional Ecology}, \emph{36}(11), 2774--2790. \url{https://doi.org/10.1111/1365-2435.14166}

\leavevmode\vadjust pre{\hypertarget{ref-lloyd1992}{}}%
Lloyd, D. G., \& Schoen, D. J. (1992). Self-and cross-fertilization in plants. I. Functional dimensions. \emph{International Journal of Plant Sciences}, \emph{153}(3, Part 1), 358--369. \url{https://doi.org/10.1086/297040}

@@

\leavevmode\vadjust pre{\hypertarget{ref-ollerton2009}{}}%
Ollerton, J., Alarcón, R., Waser, N. M., Price, M. V., Watts, S., Cranmer, L., Hingston, A., Peter, C. I., \& Rotenberry, J. (2009). A global test of the pollination syndrome hypothesis. \emph{Annals of Botany}, \emph{103}(9), 1471--1480.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-tong2016}{}}%
Tong, Z.-Y., \& Huang, S.-Q. (2016). Pre-and post-pollination interaction between six co-flowering pedicularis species via heterospecific pollen transfer. \emph{New Phytologist}, \emph{211}(4), 1452--1461. \url{https://doi.org/10.1111/nph.14005}

@@

\leavevmode\vadjust pre{\hypertarget{ref-waser1996}{}}%
Waser, N. M., Chittka, L., Price, M. V., Williams, N. M., \& Ollerton, J. (1996). Generalization in pollination systems, and why it matters. \emph{Ecology}, \emph{77}(4), 1043--1060. \url{https://doi.org/10.2307/2265575}

\leavevmode\vadjust pre{\hypertarget{ref-wilson2004}{}}%
Wilson, P., Castellanos, M. C., Hogue, J. N., Thomson, J. D., \& Armbruster, W. S. (2004). A multivariate search for pollination syndromes among penstemons. \emph{Oikos}, \emph{104}(2), 345--361. \url{https://doi.org/10.1111/j.0030-1299.2004.12819.x}
